---
title: "A Functional Approach to (Parallelised) Monte Carlo Simulation"
subtitle: "Advanced R for Econometricians"
type: "Final Project"
author: "Alexander Langnau, Öcal Kaptan, Sunyoung Ji"
discipline: "M.Sc. Econometircs"
date: "today"
studid: "232907, 230914, 229979"
supervisor: "Prof. Dr. Christoph Hanck"
secondsupervisor: "M.Sc. Martin C. Arnold, M.Sc. Jens Klenke"
ssemester: "1"
estdegree_emester: "Summer Term 2022"
deadline: "09. 09. 2022"
output:
  pdf_document:
    keep_tex: yes
    template: template.tex
    fig_caption: yes
    citation_package: biblatex
    number_sections: true
toc: true
lot: true
lof: true
graphics: true
biblio-title: References
fontsize: 11pt
geometry: lmargin=2.5cm,rmargin=2.5cm,tmargin=2.5cm,bmargin=2.5cm
biblio-files: references.bib
classoption: a4paper
language: english
---

<!-- % Template Version 1.2 -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
set.seed(123)
```

```{r library, include=FALSE}
require(utils)
library(tidyverse)
library(purrr)
```

# Introduction

Monte Carlo, named after a casino in Monaco, simulates complex probabilistic events using simple random events, such as the tossing of a pair of dice to simulate the casino’s overall business model. In Monte Carlo computing, a pseudo-random number generator is repeatedly called which returns a real number in [0, 1], and the results are used to generate a distribution of samples that is a fair representation of the target probability distribution under study. [@Barbu, Adrian_2022]
Monte Carlo Method is combined with programming in modern research and contributes to various studies.

Monte Carlo simulations are and will stay an important method in the tool box of any
econometrician, statistican or data scientist. Since these simulations  may be needed on a regular basis or are run over a complex set of functions and parameters, its time well spend to implement some
tools, that allow the user to easily create a variety of different Monte Carlo studies. 

This paper was the final project of the course "Advanced R for econometricians" at the chair of econometrics at university Duisburg Essen. The goal is to use a functional programming aproach to create a collection of different wrapper functions in R, that 

- providing a convenient interface for Monte Carlo Simulations
- create a paramter grid
- iterate homogenous function calls over the parameter grid
- provides an informative summary of the simulation results
- can be visualized by ggplot-methods
- offers the possibility to use parallelised processing (using `furrr` package)

A functional programming approach is well suited to implement the different steps.
The structure of this paper underlying code in general follows this approach:

In chapter xyz we introduce different functions, that each specifically solve the task of the bullet points mentioned above. In the beginning we´ll underline the motivation and problem behind each function and showcase the code.

At the end of each section we provide a minimal working example, that illustrates the function and its output. We tried to implement in a way, that the function works for as much cases, as possible. If there are some restrictions regarding the usage of those functions, we´ll briefly discuss them as well.

# Preprocess / Helper functions

## Function for creating grid  

`create_grid` is the function hat creates a parameter grid with all permutations of the given parameters. This is necessary to try all possible combinations to find the optimal parameters. This function tunes parameters to improve performance of Monte Carlo Simulation function.

```{r create grid}
create_grid <- function(parameters, nrep){
  input <- parameters
  storage <- list()
  name_vec <- c()
  
  for(i in 1:length(input)){ #1:3
    a <- as.numeric(input[[i]][[2]])
    b <- as.numeric(input[[i]][[3]])
    c <- as.numeric(input[[i]][[4]])
    output <- seq(from=a, to=b, by=c)
    storage[[i]] <-  output
    name_vec[i] <- input[[i]][[1]]
  }
  
  grid <- expand_grid(unlist(storage[1])
                      , unlist(storage[2])
                      , unlist(storage[3])
                      , unlist(storage[4])
                      , unlist(storage[5])
                      , c(1:nrep))
  
  names(grid) <- c(name_vec, "rep")
  
  return(grid)
}
```

`create_grid()` Example:
```{r testing create_grid}

#One parameter (works)
param_list1 <- list(c("n", 10, 20, 10))
tail(create_grid(param_list1, nrep=10), 2)
tail(create_grid(param_list1, nrep=1), 2)

#two parameter (works)
param_list2 <- list(c("n", 10, 20, 10)
                    ,c("mu", 0, 1, 0.25))
tail(create_grid(param_list1, nrep=10), 2)

#three parameters (works)
param_list3 <- list(c("n", 10, 20, 10)
                    ,c("mu", 0, 1, 0.25)
                    ,c("sd", 0, 0.3, 0.1))
tail(create_grid(param_list3, nrep=10), 2)



#four parameters (works)
param_list4 <- list(c("n", 10, 20, 10)
                    ,c("mu", 0, 1, 0.25)
                    ,c("sd", 0, 0.3, 0.1)
                    ,c("gender", 0, 1, 1))

tail(create_grid(param_list4, nrep=5),2)

grid_4 <- create_grid(param_list4, nrep=50)
tail(grid_4,2)
```

## Data generation function

`data_generation` allows users to flexibly change data while keeping the summary statistics and to choose the number of inputs by using different `purrr` mapping functions: map, map2, and pmap for a input, two inputs, and p inputs respectively.

In the function below, `simulation` means a distribution of data, and `grid` is a list of parameters. 

```{r data generate}
data_generation <- function(simulation, grid){ 
  #this is for use inside the function
  
  if(ncol(grid)==2){
    var1 <- c(unlist(grid[,1]))
    data <- map(var1, simulation) 
    #different purrr-functions depending on how many input variables we use
  }
  
  if(ncol(grid)==3){
    var1 <- c(unlist(grid[,1]))
    var2 <- c(unlist(grid[,2]))
    data <- map2(var1, var2, simulation)
  } 
  
  if(ncol(grid)==4){ 
    var1 <- c(unlist(grid[,1]))
    var2 <- c(unlist(grid[,2]))
    var3 <- c(unlist(grid[,3]))
    list1 <- list(var1,var2,var3)
    data <- pmap(list1, .f=simulation)
  } 
  
  return(data)
}
```

`data_generation()` Example:
```{r test gb}
grid1 <- create_grid(param_list1, nrep=3)
tail(data_generation(simulation=rnorm, grid=grid1),1)

grid2 <- create_grid(param_list2, nrep=3)
tail(data_generation(simulation=rnorm, grid=grid2),1)
 
```

Users can apply many distributions such as normal, uniform, poisson distributions by putting existing functions in r as `simulation`.

```{r, dg application}
# Application to Uniform distribution
param_list_runif <- list(c("n", 10, 30, 10)
                         ,c("min", 0, 0, 0)
                         ,c("max", 1, 1, 0))


grid_unif <- create_grid(param_list_runif, nrep=3)
tail(data_generation(simulation=runif, grid=grid_unif),1)

# Application to Poisson distribution

param_list_rpois <- list(c("n", 10, 30, 10)
                         , c("lambda", 0, 10, 1))

grid_pois <- create_grid(param_list_rpois, nrep=3)
tail(grid_pois,2) # nrow(grid_pois) = 99
tail(data_generation(simulation=rpois, grid=grid_pois),1)
```

## Summary function

`summary_function` offers summary statistics that users can choose.

```{r summary}
#summary function for one input
summary_function <- function(sum_fun, data_input){
  
  count <- length(data_input)
  summary_matrix <- matrix(nrow=count, ncol=1)
  
  for(i in 1:count){
    input <- list(data_input[[i]])
    output <- sapply(sum_fun, do.call, input)
    summary_matrix[i] <- output
  }
  #output <- as.data.frame(summary_matrix)
  #names(output) <- sum_fun
  colnames(summary_matrix) <- sum_fun
  return(summary_matrix)
}
```

`summary_function` Example:
```{r summary test}
grid_test <- create_grid(param_list3, nrep=3)
test_data <- data_generation(simulation=rnorm, grid=grid_test)
tail(summary_function(sum_fun=list("mean"), data_input=test_data),2)
```

## Summary array funcation

The outcome of `create_array_function` illustrates the combination of user defined grid and the summary statistics. This function product dataframes with all permutations and results that allow, thus users can look any possible parameter regarding specific grid.

```{r, array}
create_array_function <- function(comb, parameters, nrep){
  storage <- list()
  name_vec <- c()
  
  for(i in 1:length(parameters)){ 
    #this creates the sequences of parameters
    a <- as.numeric(parameters[[i]][[2]])
    b <- as.numeric(parameters[[i]][[3]])
    c <- as.numeric(parameters[[i]][[4]])
    output <- seq(from=a, to=b, by=c)
    storage[[i]] <-  output
    name_vec[i] <- parameters[[i]][[1]] 
    #this just stores the names of the variables
  }
  
  
  matrix.numeration <-  paste("rep","=", 1:nrep, sep = "")
  
  if(length(parameters)==1){
    comb_ordered <-  comb %>% arrange(comb[,2])
    seq1 <- c(unlist(storage[1]))
    
    row.names <- paste(name_vec[1],"=",seq1, sep = "")
    
    dimension_array <- c(length(seq1), nrep)
    dim_names_list <- list(row.names, matrix.numeration)
  }
  
  if(length(parameters)==2){
    comb_ordered <-  comb %>% arrange(comb[,2])  %>% arrange(comb[,3])
    seq1 <- c(unlist(storage[1]))
    seq2 <- c(unlist(storage[2]))
    
    row.names <- paste(name_vec[1],"=",seq1, sep = "")
    column.names <-  paste(name_vec[2],"=",seq2, sep = "")
    
    dimension_array <- c(length(seq1), length(seq2), nrep)
    dim_names_list <- list(row.names, column.names, matrix.numeration)
  }
  
  if(length(parameters)==3){
    comb_ordered <-  comb %>% arrange(comb[,2])  %>% 
      arrange(comb[,3]) %>% arrange(comb[,4]) 
    seq1 <- c(unlist(storage[1]))
    seq2 <- c(unlist(storage[2]))
    seq3 <- c(unlist(storage[3]))
    
    row.names <- paste(name_vec[1],"=",seq1, sep = "")
    column.names <-  paste(name_vec[2],"=",seq2, sep = "")
    matrix.names1 <-  paste(name_vec[3],"=",seq3, sep = "")
    
    dimension_array <- c(length(seq1), length(seq2), length(seq3), nrep)
    dim_names_list <- list(row.names, column.names, 
                           matrix.names1, matrix.numeration)
    
  }
  
  
  array1 <- array(comb_ordered[,ncol(comb)] 
                  #change to automatically adjust dim
                  , dim = dimension_array
                  , dim_names_list)
  return(array1)
}

```

`create_array_function` Example:
```{r, array ex}
# PREP TEST `create_array_function`
main_function_array_test <-  function(parameters #list of parameters
                                      , nrep #number of repetitions
                                      , simulation #data genereation
                                      , sum_fun){ #summary statistics
  
  grid <- create_grid(parameters, nrep) #Step 1: create grid
  raw_data <- data_generation(simulation, grid) #Step 2: simlate data
  summary <- summary_function(sum_fun, data_input=raw_data) #Step 3: Summary statistics
  comb <- cbind(grid, summary) #Step 4: Combine resuluts with parameters
  array_1 <- create_array_function(comb, parameters, nrep) #Step 5: Create array
  
  return(comb)
}

param_list3x <- list(c("n", 10, 20, 10)
                     ,c("mu", 0, 5, 1)
                     ,c("sd", 0, 1, 1))

comb1 <- main_function_array_test(parameters=param_list3x
                                  , nrep = 1
                                  , simulation = rnorm
                                  , sum_fun="mean")

head(comb1,2)

create_array_function(comb=comb1, parameters=param_list3x, nrep=1)
```




























# Monte Carlo Simulation Funcion




# Examples




# Conclusion

The above section illustrates the power of our implemented model and gives the fairly easy 
to use tool, that still allows for a variety of different specifications in terms of used parameters,
data generation processes and summary functions. Researchers, who use Monte Carlo studys on a regular basis, may save a lot of time using a tool like this in the long run.

By nature, there may be cases, where our implementation doesnt satisfy the needs of the user to the fullest, but for a wide variety of examples we showed, that it worked well and served the goal that we aimed for. Our functional programming approach allows for easy and flexible adjustments in case the use of our functions should be expanded, f.e. if a grid of more than 3 (or 4?) parameters is needed.

Theoretically, this work could be implemented as an R package to share it with the R community. But since the `MonteCarlo()` function of the `vigniette` package already provides a well working alternative to our
project besides some minor differences, there is currently no need in doing that.

















\pagebreak




